---
title: "01_Data_Wrangling"
output: html_document
date: "2024-07-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aim

The aim of this exercise is to reformat all data into a common/standardized format; and to determine if results are above PAAS levels. 


## Data sets

There are seven sources of data (included in the ./data sub folder of this project). The data sources are explained following.

* paas_lkp.csv (this contains the elemental concentrations of Post Archean Australian Shale (PAAS) which represent the background crustal abundance of elements. This is used to determine if elemental concentrations are above 'crustal abundance')

* C29030 Appendix A.xlsx (this contains data from ACARP project C29030 (Hodgkinson and Grigorescu, 2021). Each tab of this spreadsheet contains data associated with various mine sites or plant locations, and the format of data is slightly different across each tab.

* WS_REE_2023_2024.xlsx (this contains confidential data from a project area).

* BR22046077.csv & BR22046077_sample_details.csv (this contains confidential data from a project area and associated additional details)

* BR23093631.csv & BR23093631_sample_details.csv (this contains confidential data from a project area and associated additional details)

## Reformating data

The initial goal of this exercise is to reformat, standardise and merge all data sets. Further we will calculate the PAAS ratio for each sample.

The Final Data Set Should have the following column headers

"Project_Name","Sample_ID","Element_Symbol","Element_Value_ppm","Element_Description","PAAS_value_ppm","PAAS_normalised_value","Above_PASS_flag"


### Reformat C29030 Appendix A Data - Example

The following is an example of reformatting the first tab of the C29030 Appendix A.xlsx data set (named 'Collinsville'). 

In this example we will:

1. Load the data

2. Reformat the data from 'wide' to long format and clean up data

* In this step, set any 'nd' data values to -999

3. Calculate total REE values including:

  * Total_REE (Total Rare Earth Elements, which is the sum of elements La+Ce+Pr+Nd+Sm+Eu+Gd+Tb+Dy+Ho+Er+Tm+Yb+Lu, excluding Y)
  
  * Total_REE_Y (Total Rare Earth Elements, which is the sum of elements La+Ce+Pr+Nd+Sm+Eu+Gd+Tb+Dy+Ho+Er+Tm+Yb+Lu+Y)

4. Join the PASS values to the data set and add additional column including:

  * PAAS_ppm (the PAAS value for each element)

  * PAAS_normalised_value (The analysed elemental value normalised to PASS)

  * Above_PASS_flag (A flag indicating if the PAAS_normalised_value is greater than or less than the PAAS_ppm)

5. Rename columns to a consistent format

```{r}
library(tidyverse)
library(data.table)
library(openxlsx)
library(stringr)

#1. Load the data (read in the "Collinsville" tab)
d <- openxlsx::read.xlsx("./data/C29030 Appendix A.xlsx",sheet=c("Collinsville"), startRow =1,colNames=TRUE, cols=c(1:8), rows = 1:50)

#2. Reformat the data from 'wide' to long format and clean up data
m <-  d %>%  pivot_longer(cols = c(-Element,-Symbol), names_to="Sample", values_to="value") %>%
  mutate(value = str_squish(value)) %>% #get rid of white space values
  mutate(value =ifelse(value == "nd",-999,value)) %>% #set nd values to -999
  mutate(value = round(as.numeric(value),4)) %>% # round values to 4 decimal places
  na.omit() %>% #remove missing data
  mutate(Project_Area = "Collinsville") #add project area name

#3. Calculate REE values



ree <-  dcast(setDT(m),Project_Area+Sample~Symbol, value.var = "value",fill=NA) #convert back to wide format for summation
ree_calc <- ree %>% 
  mutate(REE = La+Ce+Pr+Nd+Sm+Eu+Gd+Tb+Dy+Ho+Er+Tm+Yb+Lu) %>% 
  mutate(REEY = La+Ce+Pr+Nd+Sm+Eu+Gd+Tb+Dy+Ho+Er+Tm+Yb+Lu+Y) %>% 
  mutate(LREE = La+Ce+Pr+Nd+Sm) %>% # note Pm not tested
  mutate(MREE = Eu+Gd+Tb+Dy+Y) %>% 
  mutate(HREE = Ho+Er+Tm+Yb+Lu) 

#convert back to long format so we can join PAAS values
ree_long <- melt(ree_calc,id.vars = 1:2,variable.name = "Symbol")

#4 join PASS data
paas <- read_csv("./data/paas_lkp.csv") #read paas data
ree_long <- ree_long %>% 
  inner_join(paas, by="Symbol") %>% #join paas data
  mutate(PAAS_normalised_value = value/PAAS_ppm) %>% #calculate paas normalised value 
   mutate(Above_PASS_flag = ifelse(PAAS_normalised_value >1, "Enriched above background", "Below background")) #add a flag

#5 rename columns to make consistent
names(ree_long) <- c("Project_Name","Sample_ID","Element_Symbol","Element_Value_ppm","Element_Description","PAAS_value_ppm","PAAS_normalised_value","Above_PASS_flag")

```

## Task 1

Using the above example (or your own derived version):

1. Reformat all of the data from all tabs in the "./data/C29030 Appendix A.xlsx" sheet. Tip, you might need to download and look at the excel file to see the format of the data in each tab. Note, do not bring in the calculated columns (eg descriptive statistics or PAAS calculations. Only being in the 'sample' data)

2. Merge all of the reformatted data into a single table and save the table to an .rds file in the results folder e.g: 

saveRDS('./results/C29030_data_reformatted.rds')

```{r}

# Get all sheet names (excluding non-area, Oaky Creek, and Fort Cooper sheets)
sheet_names <- setdiff(getSheetNames("./data/C29030 Appendix A.xlsx"), c("README", "Sample coordinates", "Oaky Creek", "Fort Cooper - CSIRO data"))
# nb: "Oaky Creek" and "Fort Cooper" sheets has its own pattern, so separate loading process will be done

## Initialize a list to store data from applicable sheets
sheets_data <- list()

## Loop over each sheet
for(sheet in sheet_names) {
  
  ### Read the data
  sheet_data <- read.xlsx("./data/C29030 Appendix A.xlsx", sheet = sheet, startRow = 1, colNames = TRUE, rows = 1:50)
  
  ### Getting Index of "Min" column
  min_col_position <- which(names(sheet_data) == "Min")
  
  ### Filter the column to include columns before "Min" column
  sheet_data <- sheet_data[,1:(min_col_position-1)] %>%
                pivot_longer(cols = c(-Element,-Symbol), names_to="Sample", values_to="value") %>%
                mutate(value = str_squish(value)) %>% #get rid of white space values
                mutate(value =ifelse(value == "nd",-999,value)) %>% #set nd values to -999
                mutate(value = round(as.numeric(value),4)) %>% # round values to 4 decimal places
                na.omit() %>% #remove missing data
                mutate(Project_Area = sheet) #add project area name
  
  ### Store the data in the list
  sheets_data[[sheet]] <- sheet_data
}


## Oaky Creek loading process
### Loading & Wrangling the Data 
oc_sheet_data <- read.xlsx("./data/C29030 Appendix A.xlsx", sheet = "Oaky Creek", startRow = 1, colNames = TRUE, cols=c(1:3), rows = 1:50) %>%
  pivot_longer(cols = c(-Element,-Symbol), names_to="Sample", values_to="value") %>%
  mutate(value = str_squish(value)) %>% #get rid of white space values
  mutate(value =ifelse(value == "nd",-999,value)) %>% #set nd values to -999
  mutate(value = round(as.numeric(value),4)) %>% # round values to 4 decimal places
  na.omit() %>% #remove missing data
  mutate(Project_Area = "Oaky Creek") #add project area name


## Fort Cooper loading process
### Data Loading
fc_sheet_data_raw <- read.xlsx("./data/C29030 Appendix A.xlsx", sheet = "Fort Cooper - CSIRO data", startRow = 1, colNames = TRUE)

### Data Wrangling
fc_sheet_data <- fc_sheet_data_raw %>%
  select(-c("Longitude", "Latitude")) %>%
  rename("Sample" = "Sample/Element") %>%
  mutate(across(everything(), str_squish)) %>% #get rid of white space values
  mutate(across(everything(), ~ ifelse(. == "nd", -999, .))) %>% #set nd values to -999
  mutate(across(-Sample, ~ round(as.numeric(.), 4))) %>% # round values to 4 decimal places
  pivot_longer(cols = c(-"Sample"), names_to = "Element", values_to = "value") %>%
  na.omit() %>% #remove missing data
  mutate(Project_Area = "Fort Cooper") %>% #add project area name
  left_join(select(oc_sheet_data, Element, Symbol), by="Element") %>% #join Oaky Creek data to extract "Symbol"
  mutate(Symbol = case_when(Element == "Tungsten" ~ "W", # Adding element description
                            TRUE ~ Symbol)) %>%
  select(Element, Symbol, Sample, value, Project_Area) #reorder the column

# Combine All sheets
## Combine all sheets that use for loop process
combined_df <- bind_rows(sheets_data, .id = "source") %>% select(-source)

## Combine all sheets (final combination)
C29030_data_reformatted <- bind_rows(combined_df, oc_sheet_data, fc_sheet_data)

# Saving the data into .RDS format
saveRDS(C29030_data_reformatted, './results/C29030_data_reformatted.rds')
```

## Task 2

Using the knowledge gained in Task 1; load and reformat data from BR22046077.csv into the exact same format as Task 1. 

  * For the project name use "Confidential"
  
  * After the data is reformatted, save the data as BR22046077_data_reformatted.rds in the results folder
  
    * Note that in this data set; instead of 'nd' for not detected, values will be listed as below detection with a less than sign (e.g <5). Make sure that you set any data that contains < to -999

```{r}
#here is a quick head start. the data starts on line 10 and we should use headings from line 8

BR22046077_raw <- read.csv("./data/BR22046077.csv",skip=9, col.names = names(read.csv("./data/BR22046077.csv",skip=8))) #read data (that starts on line 10) and use data from line 8 as the column names

BR22046077_data_reformatted <- BR22046077_raw %>%
  rename("Sample" = "SAMPLE") %>% # rename for consistency
  mutate(across(everything(), ~ ifelse(grepl("<", .), -999, .))) %>% #set any values that has "<" to -999
  mutate(across(everything(), str_squish)) %>% #get rid of white space values
  mutate(across(-Sample, ~ round(as.numeric(.), 4))) %>% # round values to 4 decimal places
  pivot_longer(cols = -"Sample", names_to = "Symbol", values_to = "value") %>%
  na.omit() %>% #remove missing data
  mutate(Project_Area = "Confidential") %>% #add project area name
  left_join(select(oc_sheet_data, Element, Symbol), by="Symbol") %>% #join Oaky Creek data to extract "Symbol"
  mutate(Element = case_when(Symbol == "Hf" ~ "Hafnium", #adding element description
                             Symbol == "W" ~ "Tungsten",
                             Symbol == "As" ~ "Arsenic",
                             TRUE ~ Element)) %>%
  select(Element, Symbol, Sample, value, Project_Area) #reorder the column

# Saving the data into .RDS format
saveRDS(BR22046077_data_reformatted, './results/BR22046077_data_reformatted.rds')
```


## Task 3

Load and reformat data from BR23093631.csv into the exact same format as Task 1.
  
  * For the project name use "Confidential"
  
  * After the data is reformatted, save the data as BR23093631_data_reformatted.rds in the results folder
  
  * Note that in this data set; instead of 'nd' for not detected, values will be listed as below detection with a less than sign (e.g <5). Make sure that you set any data that contains < to -999

```{r}
# 32 elements are having multiple concentration result (based on method ME-MS81 and ME-MS61r). 
# After thorough assessment, we decided that 10 elements are going to use result from ME-MS81, while the rest are going to use result from ME-MS61r. 
# Those 10 elements are Ce, Cr, Ga, Rb, Sc, Th, Ti, U, V, and W. 
# The decision was based on the concentration test, where the smaller concentration range is selected.

# Raw Data Preparation
header1 <- read.csv("./data/BR23093631.csv", header = FALSE, nrows = 1, skip = 7)  # Read first header row
header2 <- read.csv("./data/BR23093631.csv", header = FALSE, nrows = 1, skip = 8)  # Read second header row
BR23093631_raw <- read.csv("./data/BR23093631.csv", header = TRUE, skip=9) # read data (that starts on line 10)
colnames(BR23093631_raw) <- paste(header1, header2, sep = "_")  # Combine the two headers with an underscore
colnames(BR23093631_raw)[1] <- "Sample"  # Fix the first column name for consistency
BR23093631_raw <- BR23093631_raw[, -2] # remove pulp.wt column

# list of elements to be included
for_filter <- read.csv("./data/Task3_Helper.csv", sep=';')

# reformatting the table
BR23093631_data_reformatted <- BR23093631_raw %>%
  mutate(across(everything(), ~ ifelse(grepl("<", .), -999, .))) %>% #set any values that has "<" to -999
  mutate(across(everything(), str_squish)) %>% #get rid of white space values
  mutate(across(-Sample, ~ round(as.numeric(.), 4))) %>% # round values to 4 decimal places
  pivot_longer(cols = -"Sample", names_to = "sym_method", values_to = "value") %>%
  mutate(Symbol = ifelse(grepl("_", str_sub(sym_method,start=-2)), str_sub(sym_method,start=-1), str_sub(sym_method,start=-2))) %>% # extract string for symbol
  inner_join(select(for_filter, sym_method), by="sym_method") %>% # join with helper data
  na.omit() %>% #remove missing data
  mutate(Project_Area = "Confidential") %>% #add project area name
  left_join(select(oc_sheet_data, Element, Symbol), by="Symbol") %>%
  mutate(Element = case_when(Symbol == "Hf" ~ "Hafnium", #adding element description
                             Symbol == "W" ~ "Tungsten",
                             Symbol == "As" ~ "Arsenic",
                             Symbol == "Ti" ~ "Titanium",
                             TRUE ~ Element)) %>%
  select(Element, Symbol, Sample, value, Project_Area) #reorder the column
```

## Task 4

Load and reformat data from WS_REE_2023_2024.xlsx into the exact same format as Task 1.

  * note that there are two instances of the element Sc (one analysed using the ME-MS81 method, and one using the ME_4ACD81 method. In this case use the Sc value from ME-MS81 and discard the other value)
  
  * For the project name use "Confidential"
  
  * After the data is reformatted, save the data as WS_REE_2023_2024.rds in the results folder
  
  * Note that in this data set; instead of 'nd' for not detected, values will be listed as below detection with a less than sign (e.g <5). Make sure that you set any data that contains < to -999
  
```{r}
# load the raw data
ws_ree <- read.xlsx("./data/WS_REE_2023_2024.xlsx", startRow = 2, colNames = TRUE, cols=c(13:57), rows = 3:119)

# remove first row (which only contain measurement unit) and remove column 43 (because we do not use Sc from ME-4ACD81 method) from raw data
ws_ree <- ws_ree[-1,-43] 
  
# Data Wrangling
ws_ree_data_reformatted <- ws_ree %>% 
  rename("Sample" = "SAMPLE") %>%
  mutate(across(everything(), ~ ifelse(grepl("<", .), -999, .))) %>% #set any values that has "<" to -999
  mutate(across(-Sample, ~ round(as.numeric(.), 4))) %>% # round values to 4 decimal places
  pivot_longer(cols = -"Sample", names_to = "Symbol", values_to = "value") %>%
  na.omit() %>% #remove missing data
  mutate(Project_Area = "Confidential") %>% # add project area name
  left_join(select(oc_sheet_data, Element, Symbol), by="Symbol") %>% # join Oaky Creek data to extract "Symbol"
  mutate(Element = case_when(Symbol == "Hf" ~ "Hafnium", #adding element description
                             Symbol == "W" ~ "Tungsten",
                             Symbol == "As" ~ "Arsenic",
                             Symbol == "Ti" ~ "Titanium",
                             TRUE ~ Element)) %>%
  select(Element, Symbol, Sample, value, Project_Area) #reorder the column

# Saving the data into .RDS format
saveRDS(ws_ree_data_reformatted, './results/WS_REE_2023_2024.rds')
```

## Task 5

Load and merge all data sets from Task 1 to task 4 into a single object. Save the object as "all_data.rds" in the results folder

```{r}
# combine all reformatted data (from task 1 to task 4)
all_data_bfr_paas <- bind_rows(C29030_data_reformatted, BR22046077_data_reformatted, BR23093631_data_reformatted, ws_ree_data_reformatted)

# Adding Rare Earth Element (REE) Calculations
ree_all <-  dcast(setDT(all_data_bfr_paas),Project_Area+Sample~Symbol, value.var = "value",fill=NA) #convert back to wide format for summation
ree_calc_all <- ree_all %>% 
  mutate(REE = La+Ce+Pr+Nd+Sm+Eu+Gd+Tb+Dy+Ho+Er+Tm+Yb+Lu) %>% 
  mutate(REEY = La+Ce+Pr+Nd+Sm+Eu+Gd+Tb+Dy+Ho+Er+Tm+Yb+Lu+Y) %>% 
  mutate(LREE = La+Ce+Pr+Nd+Sm) %>% # note Pm not tested
  mutate(MREE = Eu+Gd+Tb+Dy+Y) %>% 
  mutate(HREE = Ho+Er+Tm+Yb+Lu) 

# Convert back to long format so we can join PAAS values
ree_long_all <- melt(ree_calc_all,id.vars = 1:2,variable.name = "Symbol")

# Join PASS data
paas <- read_csv("./data/paas_lkp.csv") #read paas data
ree_long_all <- ree_long_all %>% 
  inner_join(paas, by="Symbol") %>% #join paas data
  mutate(PAAS_normalised_value = value/PAAS_ppm) %>% #calculate paas normalised value 
  mutate(Above_PASS_flag = ifelse(PAAS_normalised_value >1, "Enriched above background", "Below background")) #add a flag

# Rename columns for consistency
names(ree_long_all) <- c("Project_Name","Sample_ID","Element_Symbol","Element_Value_ppm","Element_Description","PAAS_value_ppm","PAAS_normalised_value","Above_PASS_flag")

# Saving the data into .RDS format
saveRDS(ree_long_all, './results/all_data.rds')
```


